\Title

### 2013-01-28 synmut.tex
new figure from simulations, broader

The first result of the simulations is that genetic draft can indeed bring weakly
deleterious mutations to high frequencies and result in a dependence of the
fixation probability on initial frequency that is compatible with observations
(see \FIG{simfixpvar}).
The fixation probability of synonymous alleles decreases from the neutral
expectation to zero as their fitness effect worsens (\FIG{simfixpvar},
from $-0.0003$ in blue to $-0.003$ in red). Contextually,
the number of alleles that reach high frequencies, which is a proxy for
synonymous diversity, is reduced as well (\FIG{simfixpvar}, top inset).

To assess the reduction in fixation probability quantitatively, we calculathe
the area between the measured fixation probability and the diagonal, which is
the neutral expectation (\FIG{simfixpvar}, lower inset). This area spans a range
between $-0.5$ (no fixation at all), zero (neutral-like curve), and $+0.5$ (always
fixed). The HIV data correspond to an area of approximately $-0.2$ for synonymous
changes, and essentially zero for nonsynonymous changes.

Among all simulations, we select the one that show a large depression in
fixation probability of synonymous mutations but, simulteneously, a high
synonymous diversity, as observed in HIV data. This selection imposes a strong
constraint on the parameters, in particular on the ones related to synonymous
changes, i.e. their deleterious fraction and effect size (see \FIG{simsfig}).
A high fraction ($\gtrsim 0.8$) of only slightly deleterious synonymous sites,
with effect size $|s_d| \lesssim 0.002$, is required. This result fits well
the expectation based on the fixation/extinction times above (see \FIG{fixp1}).
Moreover, it corresponds to {\it a priori} theoretical expectations:
(a) since high frequency bins are enriched in neutral mutations (it is easier for
them to hitchhike), the hallmark of deleterious mutations will be visible there
only if they are pervasive; (b) genetic draft stops working if the
deleterious effect size becomes comparable to the adaptive effect size of escape
mutations, because the double mutant has little or no fitness advantage.

>>>>

The first result of the simulations is that genetic draft can indeed bring weakly
deleterious mutations to high frequencies and result in a dependence of the
fixation probability on initial frequency that is compatible with observations
(see \FIG{simfixpvar}).
The fixation probability of synonymous alleles decreases from the neutral
expectation to zero as their fitness effect worsens (\FIG{simfixpvar},
from $-0.0003$ in blue to $-0.003$ in red). Contextually,
the number of alleles that reach high frequencies, which is a proxy for
synonymous diversity, is reduced as well (\FIG{simfixpvar}, top inset).
To assess the reduction in fixation probability quantitatively, we calculathe
the area between the measured fixation probability and the diagonal, which is
the neutral expectation (\FIG{simfixpvar}, lower inset). This area spans a range
between $-0.5$ (no fixation at all), zero (neutral-like curve), and $+0.5$ (always
fixed). The HIV data correspond to an area of approximately $-0.2$ for synonymous
changes, and essentially zero for nonsynonymous changes.

Among all simulations, we select the one that show a large depression in
fixation probability of synonymous mutations but, simulteneously, a moderately
high synonymous diversity, quantified by the probability $P_\text{interm}$ of a
derived allele to be at intermediate frequencies $0.25 < \nu < 0.75$. HIV data show
$A_\text{syn} \sim -0.2$ and $P_\text{interm} \sim 0.005$; simulations are gated
such that $A_\text{syn} < -0.15$ and $0.0025 < P_\text{interm} < 0.010$, hence
allowing an uncertainty on mutation rates or generation times of about twofold.
This selection imposes a strong
constraint on the parameters, in particular on the ones related to synonymous
changes, i.e. their deleterious fraction and effect size (see \FIG{simsfig}).
A high fraction ($\gtrsim 0.8$) of only slightly deleterious synonymous sites,
with effect size $|s_d| \sim 0.002$, is required. This result fits well
the expectation based on the fixation/extinction times above (see \FIG{fixp1}).
Moreover, it corresponds to {\it a priori} theoretical expectations:
(a) since high frequency bins are enriched in neutral mutations (it is easier for
them to hitchhike), the hallmark of deleterious mutations will be visible there
only if they are pervasive; (b) genetic draft stops working if the
deleterious effect size becomes comparable to the adaptive effect size of escape
mutations, because the double mutant has little or no fitness advantage; (c)
mutations with a deleterious effect weaker than approximately $0.001$ are
indistinguishable from neutral ones and are drafted seamlessly to any frequency.
